---
title: "Data 607 Assignment 5"
author: "Peter Goodridge"
date: "October 1, 2017"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


###Ojbectives

*Transform and tidy the airline data*

*Perform an analysis of flight delays*

###Initialization
```{r}
library(RMySQL)
library(tidyverse)
library(ggplot2)
library(pander)
library(stringr)
library(htmlTable)
```


###Create untidy dataset
```{r}
query <- 'select * from flights'
con <- dbConnect(MySQL(), user = 'guest_user', password = 'datascience', dbname = 'airports', host = 'localhost')
untidy <- dbGetQuery(con, query)
```


###Begin Cleaning

```{r}
pander(head(untidy))


rows <- nrow(untidy)
untidy$status <-  untidy$status %>% lapply(function (x) str_replace_all(x,' ', '_'))
#We want the statuses to make better column names, so they can be referenced with the "$" operator or used in a relational DB

for (i in seq (2,rows,2)){
  untidy[i,1] <- untidy[i-1,1]
}
#The airline only appears as a header.  All the null columns beneath are assumed to be the same airline.  This is common when data is displayed, but not good for storing data.
```

###Transpose data

We want the dataframe to conform to the tidy format, so the airport and count become variables, while locations become rows.
We also split the status column into separate variables, as the count of on time and delayed represent different quantities.
Finally, we want to create variables for total flights and ratio delayed for further analysis

```{r}
tidier <- untidy %>% gather( 'Location', 'count', 3:7) %>% spread('status', 'count')
tidier$Location <-  tidier$Location %>% lapply( function (x) str_to_title(str_replace_all(x,'_', ' '))) %>% unlist()
#this no longer needs to be one word and can be propcased
tidy <- tidier %>% mutate(total_flights = delayed + on_time) %>% mutate (ratio_delayed = delayed/total_flights)

pander(tidy)
```

###Exploratory analysis

```{r}
by_airline <- tidy %>% group_by(airline) %>% 
  summarize(total_delayed = sum(delayed), total_flights = sum(total_flights), ratio = sum(delayed)/sum(total_flights))
#Compare the overal ratios

pander(data.frame(by_airline))
#pander did not work with the tibble and in my opinion looks better than htmlTable

ggplot(tidy) + geom_bar(mapping = aes(x = Location, y = ratio_delayed, fill = airline), stat = "identity") +
   labs( title = "Delay Ratio by Location", y = "Proportion Delayed")
#compare the ratios by airport

ggplot(tidy) + geom_bar(mapping = aes(x = Location, y = ratio_delayed, fill = airline), stat = "identity", 
      position ="fill") + labs( title = "Relative Delay Ratios", y = "Proportion Delayed")
#the same chart as above, but with the ratio_delayed normalized.  Can be used to see if there is a difference in delay
```

###Statistical tests

We want to prove there is a statistically signficant difference in delay activity between airlines
```{r}
t.test(tidy[tidy$airline == 'AM West', "ratio_delayed"], tidy[tidy$airline == 'Alaska', "ratio_delayed"], paired = TRUE)
by_airline %>% select(2:3) %>% chisq.test()
```

Both indicate a differance at a .05 level of significance.  The chi squared test is has much more statistical power because it differentiates between many observations and few observations.  For instance, the T test does not differentiate between 1 delay in 4 flights and 2500 delays in 10,000 flights.  This can be seen through the much high p value with this data.