---
title: "Assignment 3"
output:
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Problem 3

**1.** make all items *first* *last*
```{r}
library(stringr)
raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5543642Dr. Julius Hibbert"  

name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}")) #load in vector of names
name
library(stringr)
name_it <-function (input){
  if (str_detect(input,',')){ #not all names have a comma, so we need a conditional
    chopped <- unlist(str_split(input,',')) #split string on comma
    pasted <- str_trim(paste(chopped[2], chopped[1])) #reverse the order then paste back together
    return(pasted)
    
  }else{
    return(input)
  }
}
good_names <- as.character(sapply(name,name_it))
good_names

```
**2.** Return logical vector indicating title
```{r}
grepl("[[:alpha:]]{2,}\\.",name)
```
**3.** Does the character have a middle name?
```{r}
grepl("[[:alpha:]^.]+,.+\\>.+\\>",name)
```
###Problem 4
Describe the regular expression and provide an example
**1.** Some strange way of expressing dollars (or something to do with jquery or some other web technology)
```{r}
test <- "I have 773432$ in my pocket."
str_extract(test,"[0-9]+\\$")
```
**2.** A four letter word
```{r}
test1 <- "A test"
str_extract(test1, "\\b[a-z]{1,4}\\b")
```
**3.** The name of a text file
```{r}
test2 <- "mydata.txt"
str_extract(test2, ".*?\\.txt$")
```
**4.** A date represented with slashes
```{r}
test3 <- "today is 12/12/2012"
str_extract(test3,"\\d{2}/\\d{2}/\\d{4}")
```
**5.** An html element
```{r}
test4 <- "<h1>this is my header</h1>"
str_extract(test4, "<(.+?)>.+?</\\1>")
```
###Problem 9
```{r}
long_string <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!"
message <- unlist(str_extract_all(long_string, "[[:Upper:][:punct:]]+"))#find all upper case letters and punctionation
paste(unlist(strsplit (paste(message, sep = '', collapse = ''),"\\.")),collapse = ' ')
#Paste it together then split it on "."
```